---
title: "trip-advisors-hotel-rating-prediction"
author: "Orion-lee"
date: "09/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Trip Advisor Hotel Ratings Prediction through User Reviews
 
## 1. Introduction

Moving into the digital age, booking platforms has been providing direct B2C channels and strike the hotel industries. Meanwhile, first person experiences on social medias or any popular tourist platforms, such as Trip Advisor, has accounted for a significant part in customers' perceptions during their purchasing process. However, a good customer rating does not always go align with the traditional aesthetics standard of rating the star-level of a hotel. For example, a cozy hostel at a convenient location can always be rated higher than a renowned luxury one. These kind qualitative quantities are more likely be extracted from the customer reviews instead of the general description, eg. room size, pricing, facilities, etc., of a hotel.

Adapting to the change, business owners set up digital marketing teams to better market themselves online. These teams need not only to focus on the blogs/columns of their own hotels, but also studying the industry leaders in understanding the customer preferences. Nevertheless, reading every piece of reviews is time consuming while building correlations between word descriptions and a hotel rating is seemingly impossible by manual efforts.

Therefore, we will apply our knowledge in Data Science try to predict a rating with only the user reviews and identify if any significant predators (eg. description text) marketers should look for in reading each review.

## 1.1. Data Set

Our data set is downloaded from Kaggle at https://www.kaggle.com/andrewmvd/trip-advisor-hotel-reviews provided andrewmvd, a Kaggle Grandmaster user.

```{r cars}
library(tidyverse)
url <- 'http://www.trumptwitterarchive.com/data/realdonaldtrump/%s.json'
reviews <- as.data.frame()
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
